---
title: "Finding ATTAINS Features"
format: html
editor: visual
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    number_sections: false
vignette: >
  %\VignetteIndexEntry{Finding ATTAINS Features}
  %\usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 78
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  collapse = TRUE,
  fig.width = 8,
  fig.height = 8
)
```

```{r, include = FALSE}
# load in all necessary packages
# source("src/setup.R")
#... currently: "TADA", "tidyverse", "sf", "TADA", "urltools", "geojsonsf", "leaflet", "viridis", "tigris", "readxl"
# load library for testing
library(TADA);library(knitr);library(tidyverse);library(sf);library(urltools);library(geojsonsf)
source('src/TADA_MakeSpatial.R')
source('src/TADA_GetATTAINS.R')
source('src/TADA_ViewATTAINS.R')
# a function used in ATTAINS functions, but will not be an exported function in TADA
source('src/fetchATTAINS.R')
```

This vignette represents three functions aimed towards completing the TADA
Module 2 goal: *Match WQP sites with ATTAINS assessment units. Assist user
with reviewing sites and Assessment Units (AUs) on a map and deciding to keep
or exclude sites (systematically first, maybe using catchments as a buffer
around the AU points/line/polygons for consistency, and then let users make
manual edits to include or exclude sites (main requirement)). Consider adding
option to create new AUs for new areas where WQP sites are available but no AU
exists yet (secondary requirement). Output is a simple spreadsheet linking
ATTAINS AU IDs with WQP Location IDs.*

**A Note About ATTAINS:**

The Assessment, Total Maximum Daily Load (TMDL) Tracking and Implementation
System ([ATTAINS](https://www.epa.gov/waterdata/attains)) is an online
platform that organizes and combines each state’s Clean Water Act reporting
data into a single data repository. The geospatial component of ATTAINS
includes spatial representations of each state’s assessment units as well as
their assigned designated uses, their most recent EPA reporting category
(i.e., their impairment status), their impaired designated uses, and the
parameter(s) causing the impairment.

Within an assessment unit, the water quality standards remain the same and all
water features are assessed as one entity. Depending on the state, these
assessment units can be a specific point along a river, a river reach, an
entire river, or even an entire watershed. (In other words, assessment units
can take the form of point, line, and area features, or some combination of
all of them.) Moreover, it is possible that some assessment units are not
geospatially referenced at all, meaning they are not captured in the ATTAINS
geospatial database.

## `TADA_MakeSpatial()`

This function converts any Water Quality Portal (WQP)-style data frame with
latitude/longitude data into an {sf} object and is the backbone for future
geospatial TADA functions. To run the function, the user supplies a WQP data
frame and the coordinate reference system that they want the spatial object to
be in (the default is WGS 84). For the function to work properly, the input
data frame must have - at a minimum - WQP observation coordinates in
"LongitudeMeasure" and "LatitudeMeasure" columns (or "TADA.LongitudeMeasure"
and "TADA.LatitudeMeasures" columns if the user chose to auto-clean their WQP
results) and a "HorizontalCoordinateReferenceSystemDatumName" column.

#### Using `TADA_MakeSpatial()`

First, we will need to pull in some TADA Water Quality Portal Data:

```{r}
# pH data in Larimer County, Colorado from 
TADA_dataframe <- TADA_DataRetrieval(startDate = "2020-01-01",
                                     endDate = "2020-12-31",
                                     characteristicName = "pH",
                                     countycode = "US:08:069",
                                     applyautoclean = TRUE)
```

Now, we can make the water quality data spatial by running
`TADA_MakeSpatial()`:

```{r}
# default CRS is WGS84 (4326)
TADA_spatial <- TADA_MakeSpatial(data = TADA_dataframe, crs = 4326)
```

This new spatial object is identical to the original TADA dataset, but now
includes a "geometry" column, which allows for mapping and additional
geospatial capabilities. Enter `?TADA_MakeSpatial` into the console to review
another example of this function in use and additional information.

```{r}
leaflet::leaflet() %>% 
  leaflet::addProviderTiles("Esri.WorldTopoMap", 
                            group = "World topo", 
                            options = leaflet::providerTileOptions(updateWhenZooming = FALSE,
                                                                   updateWhenIdle = TRUE)) %>% 
  leaflet::clearShapes() %>% 
  leaflet.extras::addResetMapButton()  %>%
  leaflet::addLegend(position = "bottomright",
                     colors = "black",
                     labels = "Water Quality Observation(s)",
                     opacity = 1) %>%
  leaflet::addCircleMarkers(data = TADA_spatial,
                            color = "grey", fillColor = "black", 
                            fillOpacity = 0.8, stroke = TRUE, weight = 1.5, radius = 6, 
                            popup = paste0("Site ID: ", 
                                           TADA_spatial$MonitoringLocationIdentifier, 
                                           "<br> Site Name: ",
                                           TADA_spatial$MonitoringLocationName))
```

## `TADA_GetATTAINS()`

This function pulls in ATTAINS data from the EPA's ATTAINS Assessment
Geospatial Service and links it to TADA-pulled Water Quality Portal
observations. For the function to work properly, the input data frame must
have - at a minimum - WQP observation coordinates in "LongitudeMeasure" and
"LatitudeMeasure" columns (or "TADA.LongitudeMeasure" and
"TADA.LatitudeMeasures" columns if the user chose to auto-clean their WQP
results) and a "HorizontalCoordinateReferenceSystemDatumName" column.

Users also have the option of loading the ATTAINS geospatial {sf} objects into
their local environment with `return = TRUE`. The objects returned are named
`ATTAINS_catchments`, `ATTAINS_lines`, `ATTAINS_points`, and
`ATTAINS_polygons`.

Regardless of the user's decision on returning the ATTAINS {sf} objects,
`TADA_GetATTAINS()` always returns a dataframe containing the original TADA
WQP dataset, plus new columns representing the ATTAINS assessment unit(s) that
fall within the same NHDPlus HR catchment as them. This means that it is
possible for a single TADA WQP observation to have multiple ATTAINS assessment
units linked to it and subsequently more than one row of data. Such WQP
observations can be identified using the `index` column (i.e., multiple rows
with the same index value are the same observation).

#### Using `TADA_GetATTAINS()`

Using either our original `TADA_dataframe` *or* the geospatial version
`TADA_spatial`, we can pull in the ATTAINS features that are within the same
NHD HR catchment as our observations:

```{r}
TADA_with_ATTAINS <- TADA_GetATTAINS(data = TADA_dataframe, return = TRUE)

TADA_with_ATTAINS <- TADA_GetATTAINS(data = TADA_spatial, return = TRUE)
```

This new `TADA_with_ATTAINS` object is identical to the original TADA Water
Quality Portal dataframe, but now has additional columns associated with the
ATTAINS assessment unit(s) that lie in the same NHD HR catchment as them
(these columns are prefixed with "ATTAINS"). However, because our
`TADA_with_ATTAINS` object contains more rows than the original TADA
dataframe, we can deduce that some water quality observations fall within an
NHD HR catchment that contains multiple ATTAINS assessment units.

And, because we set `return = TRUE`, we now have all the raw ATTAINS features
associated with these TADA Water Quality Portal observations saved in our
Global Environment as `ATTAINS_catchments`, `ATTAINS_lines`, `ATTAINS_points`,
and `ATTAINS_polygons` objects.

Enter `?TADA_GetATTAINS` into the console to review another example of this
function in use and additional information.

## `TADA_ViewATTAINS()`

This function visualizes the raw ATTAINS features that are linked to the TADA
WQP observations. (Essentially, this maps the same data created from running
`TADA_GetATTAINS()`.) For the function to work properly, the input data frame
must be a WQP dataframe developed by `TADA_DataRetrieval()`.

#### Using `TADA_ViewATTAINS()`

Using either our original `TADA_dataframe` *or* the geospatial version
`TADA_spatial`, we can plot all water quality observations, as well as the
ATTAINS features, using `TADA_ViewATTAINS()` . Enter `?TADA_ViewATTAINS` into
the console to review another example query and additional information.

```{r}
TADA_ViewATTAINS(data = TADA_spatial)
```

\
